Liveness-Based Garbage Collection for Lazy Languages

### 2015-11-27 paper/paper.tex
fix labels

We shall now present a proof of soundness of the analysis presented in
Section~\ref{sec:liveness-analysis}.  \warning{While  the proof  is on
  the broad lines of  \cite{asati14lgc}, there are several differences
  due to the  lazy semantics of the language in  this paper.} Here are
the ideas behind the proof.
\begin{enumerate}
\item       We        enrich       the        standard       semantics
  (Figure~\ref{fig:lang-semantics}) to model  a liveness-based garbage
  collection  just  before  each  $\rightsquigarrow$  transition.  The
  simulated garbage collection starts from  the root-set and inserts a
  special value  $\bot$ at each location  in the heap that  contains a
  reference that is reachable but not live. Any attempt to dereference
  such  locations during  the  transition will  result  in entering  a
  special state denoted \bang. We call the semantics after
  augmentation,  \emph{minefield semantics}.
\item \label{inline} Assuming that a  program enters the \bang\ state,
  we  show  how to  construct,  through  inline expansion,  a  program
  without function calls which has the same minefield behavior.
\item  The final  step   shows  that  no program  without
  function calls can enter a \bang\ state. As a consequence no program
  (with or without function calls) can enter the \bang\ state. 
\end{enumerate}

@@

We shall now  present a proof of soundness of  the liveness analysis
presented in Section~\ref{sec:liveness-analysis}.   While the proof is
on the broad lines of \cite{asati14lgc}, there are several differences
due to  the lazy  semantics of  the language  being 
analyzed in this paper. Here is the idea behind the proof.
\begin{enumerate}
\item       We        enrich       the        standard       semantics
  (Figure~\ref{fig:lang-semantics}) to model  a liveness-based garbage
  collection  just  before  each  $\rightsquigarrow$  transition.  The
  simulated garbage collection starts from  the root-set and inserts a
  special value  $\bot$ at each location  in the heap that  contains a
  reference that is reachable but not live. Any attempt to dereference
  such  locations during  the  transition will  result  in entering  a
  special state denoted \bang. We call the semantics after
  augmentation  \emph{minefield semantics}.
\item \label{inline} Assuming that a  program enters the \bang\ state,
  we  show  how to  construct,  through  inline expansion,  a  program
  without function calls which has the same minefield behavior.
\item  The final  part  of the  proof shows  that  no program  without
  function calls can enter a \bang\ state. As a consequence no program
  (with or without function calls) can enter the \bang\ state. 
\end{enumerate}

@@

\begin{lemma}
\label{lemma:call-less-cannot-go-bang}
Consider the minefield execution of  a program without function calls.
Such a program cannot enter the \bang\ state.
\end{lemma}
\begin{proof}
Consider a state in the minefield execution of a program $(\rho, S, H,
e, \sigma)$.   We show  by induction  on the  number $n$  of minefield
steps  leading to  this state  that  the next  step cannot  lead to  a
\bang\ state.

@@

\begin{proof}
Assume to  the contrary that  a program $P$  enters the
\bang\  state.  We can  transform  $P$  to a  call-less
program $P'$ such that  the minefield executions of $P$
and $P'$  are identical  except for change  of variable
names.  \warning{Check  whether  the previous  line  is
  correct.}                 However,                 by
Lemma~\ref{lemma:call-less-cannot-go-bang} we know that
$P'$,   a   call-less   program,   cannot   enter   the
\bang\  state.  Therefor  $P$  also  cannot  enter  the
\bang\ state.
\end{proof}

>>>>

We shall now present a proof of soundness of the analysis presented in
Section~\ref{sec:liveness-analysis}.  \warning{While  the proof  is on
  the broad lines of  \cite{asati14lgc}, there are several differences
  due to the  lazy semantics of the language in  this paper.} Here are
the ideas behind the proof.
\begin{enumerate}
\item       We        enrich       the        standard       semantics
  (Figure~\ref{fig:lang-semantics}) to model  a liveness-based garbage
  collection  just  before  each  $\rightsquigarrow$  transition.  The
  simulated garbage collection starts from  the root-set and inserts a
  special value  $\bot$ at each location  in the heap that  contains a
  reference that is reachable but not live. Any attempt to dereference
  such  locations during  the  transition will  result  in entering  a
  special state denoted \bang. We call the semantics after
  augmentation,  \emph{minefield semantics}.
\item \label{inline1} Assuming that a  program enters the \bang\ state,
  we  show  how to  construct,  through  inline expansion,  a  program
  without function calls which has the same minefield behavior.
\item  The final  step   shows  that  no program  without
  function calls can enter a \bang\ state. As a consequence no program
  (with or without function calls) can enter the \bang\ state. 
\end{enumerate}

@@

We shall now  present a proof of soundness of  the liveness analysis
presented in Section~\ref{sec:liveness-analysis}.   While the proof is
on the broad lines of \cite{asati14lgc}, there are several differences
due to  the lazy  semantics of  the language  being 
analyzed in this paper. Here is the idea behind the proof.
\begin{enumerate}
\item       We        enrich       the        standard       semantics
  (Figure~\ref{fig:lang-semantics}) to model  a liveness-based garbage
  collection  just  before  each  $\rightsquigarrow$  transition.  The
  simulated garbage collection starts from  the root-set and inserts a
  special value  $\bot$ at each location  in the heap that  contains a
  reference that is reachable but not live. Any attempt to dereference
  such  locations during  the  transition will  result  in entering  a
  special state denoted \bang. We call the semantics after
  augmentation  \emph{minefield semantics}.
\item \label{inline2} Assuming that a  program enters the \bang\ state,
  we  show  how to  construct,  through  inline expansion,  a  program
  without function calls which has the same minefield behavior.
\item  The final  part  of the  proof shows  that  no program  without
  function calls can enter a \bang\ state. As a consequence no program
  (with or without function calls) can enter the \bang\ state. 
\end{enumerate}

@@

\begin{lemma}
\label{lemma:call-less-cannot-go-bang-app}
Consider the minefield execution of  a program without function calls.
Such a program cannot enter the \bang\ state.
\end{lemma}
\begin{proof}
Consider a state in the minefield execution of a program $(\rho, S, H,
e, \sigma)$.   We show  by induction  on the  number $n$  of minefield
steps  leading to  this state  that  the next  step cannot  lead to  a
\bang\ state.

@@

\begin{proof}
Assume to  the contrary that  a program $P$  enters the
\bang\  state.  We can  transform  $P$  to a  call-less
program $P'$ such that  the minefield executions of $P$
and $P'$  are identical  except for change  of variable
names.  \warning{Check  whether  the previous  line  is
  correct.}                 However,                 by
Lemma~\ref{lemma:call-less-cannot-go-bang-app} we know that
$P'$,   a   call-less   program,   cannot   enter   the
\bang\  state.  Therefor  $P$  also  cannot  enter  the
\bang\ state.
\end{proof}
